---
title: "Biostats 750 Final Project"
author: "Derrick Liu"
date: "4/21/2021"
output: html_document
---

Study Objective: To determine which of the 13 predictor variables in the data 
set have a statistically significant impact on whether a person experiences a 
heart attack. 

```{r}
# reading in the data set and viewing the first few rows

heart_attack_dat <- read.csv("heart.csv")
heart_attack_dat_2 <- heart_attack_dat[-86,]
heart_attack_dat_2
```

# Data Visualization

```{r}
# Comparing relationship between Heart Attack Incidence (0/1) and Cholesterol
# Level

library(ggplot2)

ggplot(data = heart_attack_dat_2, aes(x = chol, y = output)) +
  geom_point() + 
  ggtitle("Did Heart Attack Happen vs. Cholesterol Level") + 
  xlab("Cholesterol Level") + 
  ylab("Did Heart Attack Happen")
```

160-189 mg/dl is considered high for cholesterol

cholesterol levels > 200 seem really odd --> is this data accurate? 

```{r}
# getting summary statistics for each of the numeric type predictor variables to 
# be used from the data set

library(stargazer)

stargazer(heart_attack_dat_2[c("age","trtbps","chol","thalachh","oldpeak")], type="text",summary.stat=c("n","mean","sd", "min", "max"))
```

histogram of the resting blood pressure
```{r}
hist(heart_attack_dat$trtbps, main = "Resting Blood Pressure", 
     xlab = "Resting Blood Pressure (in mm Hg)")
```

```{r}
hist(heart_attack_dat$chol, main = "Cholesterol Levels", 
     xlab = "Cholesterol (mg/dl)")
```


Initial observations from summary table above:

-The blood pressures for people in this data set are also seemingly really high. 
A healthy systolic blood pressure is between 90 and 120 mm Hg in an adult, 
according to Healthline. 
-The cholesterol levels for people in the data set were very, very high. For 
reference, a healthy level of total cholesterol is 125-200 mg/dL, according to 
Medline Plus. 

# Applying Methods That We Learned in Class

```{r}
# fitting a logistic regression model with all 13 predictors

heart_attack_logreg <- glm(output ~ .,
                           data = heart_attack_dat_2)
summary(heart_attack_logreg)
```

```{r}
# fitting a logistic regression model with the predictors that were found to be 
# statistically significant in the full model above

heart_attack_logreg2 <- glm(output ~ sex+cp+thalachh+exng+oldpeak+caa+thall,
                           data = heart_attack_dat_2)
summary(heart_attack_logreg2)
```

All of the predictors that were found to be statistically significant in the 
full model are all still statistically significant when they are included
in the reduced model.

```{r}
# running LDA

library(MASS)

lda.fit = lda(output ~ sex+cp+thalachh+exng+oldpeak+caa+thall, 
              data = heart_attack_dat)

lda.fit
summary(lda.fit)

# how to interpret LDA output
```

```{r}
# running LDA

library(MASS)

qda.fit = qda(output ~ sex+cp+thalachh+exng+oldpeak+caa+thall, 
              data = heart_attack_dat)

qda.fit
summary(qda.fit)

# how to interpret QDA output
```

```{r}
# running 1000 bootstrap replications with 200 samples

n_rows = nrow(heart_attack_dat)
n = 200 # number of samples
N = 1000 # number of bootstrap replications

for(i in 1:N){
  indices <- sample(1:n_rows, n_rows, replace = TRUE)
  
  # need to update a variable that keeps track of number of obs predicted to have heart attack
}

# can perhaps consider all the bootstrap samples as the training sample and the 
# original data set as the test sample
```

