---
title: "Final Draft"
author: "Derrick Liu"
date: "5/8/2021"
output: html_document
---

```{r}
# reading in the data set

heart_dat <- read.csv("heart.csv")
head(heart_dat)
```

# Data visualization

```{r}
# histogram of the age variable

hist(heart_dat$age, main = "Ages", xlab = "Age (in years)")
```

```{r}
# histogram of the resting blood pressure

hist(heart_dat$trtbps, main = "Resting Blood Pressure", 
     xlab = "Resting Blood Pressure (in mm Hg)")
```

```{r}
# log(trtbps+1) transformation to see if that fixes the skewness to the right
# read that log(x+1) transformation works to fix distributions skewed to the 
# right at the following link: http://seismo.berkeley.edu/~kirchner/eps_120/Toolkits/Toolkit_03.pdf 
# looks a lot more normally distributed

library(base)

heart_dat_mod <- heart_dat
heart_dat_mod$trtbps_transformed <- log10((heart_dat_mod$trtbps)+1)

hist(heart_dat_mod$trtbps_transformed, main = "Resting Blood Pressure", 
     xlab = "Resting Blood Pressure (in mm Hg)")
```

```{r}
# histogram of the cholesterol levels

hist(heart_dat$chol, main = "Cholesterol Levels", 
     xlab = "Cholesterol (mg/dl)")
```

```{r}
# log(chol+1) transformation to see if that fixes the skewness to the right
# looks a lot more normally distributed

heart_dat_mod$chol_transformed <- log10((heart_dat_mod$chol)+1)

hist(heart_dat_mod$chol_transformed, main = "Cholesterol Levels", 
     xlab = "Cholesterol (mg/dl)")
```

```{r}
# histogram of the maximum heart rate achieved
# looks like there is a skew to the left

hist(heart_dat$thalachh, main = "Maximum Heart Rate", 
     xlab = "Maximum Heart Rate Achieved (in bpm)")
```

```{r}
# sqrt(constant-thalachh) transformation to see if that fixes the skewness to the left
# looks a lot more normally distributed
# coombatting left skew: https://rcompanion.org/handbook/I_12.html#:~:text=For%20left%2Dskewed%20data%E2%80%94tail,log%20(constant%20%E2%80%93%20x).

heart_dat_mod$thalachh_transformed <- sqrt(250-(heart_dat_mod$thalachh))

hist(heart_dat_mod$thalachh_transformed, main = "Maximum Heart Rate", 
     xlab = "Maximum Heart Rate Achieved (in bpm)")
```

```{r}
# histogram of the oldpeak variable
# looks like there is a strong skew to the right

hist(heart_dat$oldpeak, main = "ST Depression", 
     xlab = "ST Depression Induced by Exercise Relative to Rest")
```

```{r}
# log(oldpeak+1) transformation to see if that fixes the skewness to the right
# the skew to the right seems to be less by a tiny bit, but it is still there

heart_dat_mod$oldpeak_transformed <- log10((heart_dat_mod$oldpeak)+1)

hist(heart_dat_mod$oldpeak_transformed, main = "ST Depression", 
     xlab = "ST Depression Induced by Exercise Relative to Rest")
```

0 = female, 1 = male, more than twice as many men in this dataset
```{r}
library(dplyr)

heart_dat %>% group_by(sex) %>% summarise( total = n())
```

```{r}
# modified data set with transformed variables to use for model in LDA/QDA
# so the normality assumption for the data is not violated

heart_dat_modified <- heart_dat_mod[,-c(4,5,8,18)]
head(heart_dat_modified)
```


correlation amongst all the predictors, non seem to be super correlated
```{r}
round(cor(heart_dat), 2)
```

```{r}
# getting summary statistics for each of the numeric type predictor variables to 
# be used from the data set

library(stargazer)

stargazer(heart_dat[c("age","trtbps","chol","thalachh","oldpeak")], type="text",summary.stat=c("n","mean","sd", "min", "max"))
```

```{r}
# changing all the non-numeric variables to factor type variables

heart_dat$sex <- as.factor(heart_dat$sex)
heart_dat$cp <- as.factor(heart_dat$cp)
heart_dat$fbs <- as.factor(heart_dat$fbs)
heart_dat$restecg <- as.factor(heart_dat$restecg)
heart_dat$exng <- as.factor(heart_dat$exng)
heart_dat$slp <- as.factor(heart_dat$slp)
heart_dat$caa <- as.factor(heart_dat$caa)
heart_dat$thall <- as.factor(heart_dat$thall)
#heart_dat$output <- as.factor(heart_dat$output)

head(heart_dat)
```

# Lasso and Ridge

```{r}
# lasso - code adapted from Lab 5
# binomial still not working here

library(glmnet)

train=sample(1:nrow(heart_dat), nrow(heart_dat)/2)
test = (-train)

grid=10^seq(10,-2,length=100)

x=model.matrix(output~.,heart_dat)[,-14]
y=heart_dat$output
y.test=y[test]

lasso.mod=glmnet(x[train,],y[train],alpha=1,lambda=grid)
#lasso.mod=glmnet(x[train,],y[train],alpha=1,lambda=grid, family = binomial)
plot(lasso.mod)
```

```{r}
# continuing lasso - code adapted from Lab 5

set.seed(1)
cv.out=cv.glmnet(x[train,],y[train],alpha=1)
#cv.out=cv.glmnet(x[train,],y[train],alpha=1, family = binomial)
plot(cv.out)

bestlam=cv.out$lambda.min # tuning parameter
lasso.pred=predict(lasso.mod,s=bestlam,newx=x[test,])
mean((lasso.pred-y.test)^2) # MSE

out = glmnet(x,y,alpha=1,lambda=grid) # producing a model with 10 predictors
lasso.coef=predict(out,type="coefficients",s=bestlam)[1:13,]
lasso.coef[lasso.coef!=0] 

bestlam
cv.out$lambda

```

```{r}
# ridge - code adapted from Lab 5

ridge.mod=glmnet(x[train,],y[train],alpha=0,lambda=grid, thresh=1e-12)
#ridge.mod=glmnet(x[train,],y[train],alpha=0,lambda=grid, thresh=1e-12, 
                 #family = binomial)
ridge.pred=predict(ridge.mod,newx=x[test,])
mean((ridge.pred-y.test)^2)

out2=glmnet(x,y,alpha=0)  # again, glmnet function is producing a model with not all zero values for the parameter estimates
# glmnet function fits a GLM with lasso or elastic net regularization
ridge.coef = predict(out2,type="coefficients",s=bestlam)[1:13,]
ridge.coef[ridge.coef!=0] 
```

# Fitting a logistic regression model

```{r}

# fitting a model with best subset selection

library(MASS) # for the stepAIC function
library(dplyr) # for the pipeline operator

set.seed(1)

best_subset_logregmodel <- glm(output ~., data = heart_dat, family = binomial) %>%
  stepAIC(trace = FALSE)

best_subset_logregmodel
summary(best_subset_logregmodel)
```

```{r}
# fitting a model with forward stepwise selection

set.seed(1)
forward_logregmodel <- glm(output ~., data = heart_dat, family = binomial) %>%
  stepAIC(direction = "forward", trace = FALSE)

forward_logregmodel
summary(forward_logregmodel)
```

```{r}
# fitting a model with backward stepwise selection

set.seed(1)
backward_logregmodel <- glm(output ~., data = heart_dat, family = binomial) %>%
  stepAIC(direction = "backward", trace = FALSE)

backward_logregmodel
summary(backward_logregmodel)
```

# Fitting our final model (for now) chosen from AIC on the original dataset

```{r}
final_logregmodel <- glm(output ~sex+cp+trtbps+exng+oldpeak+slp+caa+thall, 
                         data = heart_dat, family = binomial) 
summary(final_logregmodel)
```

# k-fold Cross Validation

```{r}
# Cross Validation done on the model chosen by best subset selection

k = 10
folds = sample(1:k, nrow(heart_dat), replace = TRUE)

set.seed(1)
model_saver <- matrix(ncol = 10)
cv.errors2 = matrix(NA, k, 13)
intercept <- matrix(NA, k, 11)
predictors_saver <- matrix(NA, k, 1)

for(j in 1:k){ 

   best_subset_logregmodel <- glm(output ~., data = heart_dat[folds != j,], family = binomial) %>%
  stepAIC(trace = F)
  model_saver[j] <- length(best_subset_logregmodel$coefficients)  #saves the number of non-zero coefficients
  intercept[j,] <- best_subset_logregmodel$coefficients[1:11]     #looks at model coefficients
  predictors <- summary(best_subset_logregmodel)
  predictors_saver[j] <- c(predictors$call)    #Saves the model predictors
  
    for(i in 1:13){
    pred=predict(best_subset_logregmodel,heart_dat[folds==j,], type = "response") 
    pred = ifelse(pred > 0.5, 1, 0)
    cv.errors2[j, i]=mean((heart_dat$output[folds==j] != pred))
    }
}

mean.cv.errors.best = mean(cv.errors2)
#mean.cv.errors.best         # gives the average misclassification rate
#intercept                 #These are the coefficents
predictors_saver         # These are the predictors and the models    
```

```{r}
# Cross Validation done on the model chosen by forward stepwise selection

cv.errors3 = matrix(NA, k, 13)
intercept2 <- matrix(NA, k, 14)
model_saver2 <- matrix(ncol = 14)
predictors2_saver <- matrix(NA, k, 1)

for(j in 1:k){ 
  
  forward_logregmodel <- glm(output ~., data = heart_dat[folds != j,], family = binomial) %>%
  stepAIC(direction = "forward", trace = FALSE)
  model_saver2[j] <- length(forward_logregmodel$coefficients)   #saves the number of non zero coefficients
  intercept2[j,] <- forward_logregmodel$coefficients[1:14] #looks at model coeffeicent
    predictors2 <- summary(forward_logregmodel)
  predictors2_saver[j] <- c(predictors2$call)
  
    for(i in 1:13){  
    pred=predict(forward_logregmodel,heart_dat[folds==j,], type = "response") 
    pred = ifelse(pred > 0.5, 1, 0)
    cv.errors3[j,i]=mean(heart_dat$output[folds==j] != pred)
    } 
}

mean.cv.errors.forward = mean(cv.errors3)
#mean.cv.errors.forward     #gives the average misclassification rate
##intercept2            #These are the coefficients
predictors2_saver        #These are the models and predictors
```

## AUC Area under the curve 
```{r}
library(ROCR)

```


backwards model
```{r}
back.preds <- predict(backward_logregmodel, data = heart_dat, type = "response")
back.preds <- ifelse(back.preds < 0.5, 0, 1)
back.rocr <- prediction(back.preds, heart_dat$output)

back.perf <- performance(back.rocr,"tpr","fpr")
plot(back.perf, avg="threshold",
     spread.estimate="boxplot")

performance(back.rocr,"auc")@y.values
```


forwards model
```{r}
forw.preds <- predict(forward_logregmodel, data = heart_dat, type = "response")
forw.preds <- ifelse(forw.preds < 0.5, 0, 1)
forw.rocr <- prediction(forw.preds, heart_dat$output)

forw.perf <- performance(forw.rocr,"tpr","fpr")
plot(forw.perf, avg="threshold",
     spread.estimate="boxplot")

performance(forw.rocr,"auc")@y.values
```


best subset model
```{r}
best.preds <- predict(best_subset_logregmodel, data = heart_dat, type = "response")
best.preds <- ifelse(best.preds < 0.5, 0, 1)
best.rocr <- prediction(best.preds, heart_dat$output)

best.perf <- performance(best.rocr,"tpr","fpr")
plot(best.perf, avg="threshold",
     spread.estimate="boxplot")

performance(best.rocr,"auc")@y.values
```


Final logreg model (same as the model chosen by best subset selection/backward stepwise selection)
```{r}
final.preds <- predict(final_logregmodel, data = heart_dat, type = "response")
final.preds <- ifelse(final.preds < 0.5, 0, 1)
final.rocr <- prediction(final.preds, heart_dat$output)

final.perf <- performance(final.rocr,"tpr","fpr")
plot(final.perf, avg="threshold",
     spread.estimate="boxplot")

performance(final.rocr,"auc")@y.values
```

# For LDA and QDA

```{r}
# changing all the non-numeric variables to factor type variables

heart_dat_modified$sex <- as.factor(heart_dat_modified$sex)
heart_dat_modified$cp <- as.factor(heart_dat_modified$cp)
heart_dat_modified$fbs <- as.factor(heart_dat_modified$fbs)
heart_dat_modified$restecg <- as.factor(heart_dat_modified$restecg)
heart_dat_modified$exng <- as.factor(heart_dat_modified$exng)
heart_dat_modified$slp <- as.factor(heart_dat_modified$slp)
heart_dat_modified$caa <- as.factor(heart_dat_modified$caa)
heart_dat_modified$thall <- as.factor(heart_dat_modified$thall)
#heart_dat$output <- as.factor(heart_dat$output)

head(heart_dat_modified)
```


```{r}

# fitting a model with best subset selection

set.seed(1)

best_subset_logregmodel2 <- glm(output ~., data = heart_dat_modified, family = binomial) %>%
  stepAIC(trace = FALSE)

best_subset_logregmodel2
summary(best_subset_logregmodel2)
```

```{r}
# fitting a model with forward stepwise selection

set.seed(1)
forward_logregmodel2 <- glm(output ~., data = heart_dat_modified, family = binomial) %>%
  stepAIC(direction = "forward", trace = FALSE)

forward_logregmodel2
summary(forward_logregmodel2)
```

```{r}
# fitting a model with backward stepwise selection

set.seed(1)
backward_logregmodel2 <- glm(output ~., data = heart_dat_modified, family = binomial) %>%
  stepAIC(direction = "backward", trace = FALSE)

backward_logregmodel2
summary(backward_logregmodel2)
```

Again, we see that the model chosen (with lowest AIC) by best subset selection
and backward stepwise selection are the same, even with the transformed variables
in place of the original variables trtbps, chol, and thalachh. 

```{r}
# LDA for the model chosen by best subset selection/backward stepwise selection

set.seed(1)
dt2 = sort(sample(nrow(heart_dat_modified), nrow(heart_dat_modified)*.9))
train2 <- heart_dat_modified[dt2,]
test2 <- heart_dat_modified[-dt2,]

lda_best <- lda(output~ sex+cp+exng+oldpeak+slp+caa+thall+trtbps_transformed+thalachh_transformed, 
                data = train2)
lda_best_pred <- predict(lda_best, test2)

table(lda_best_pred$class, test2$output)
```

```{r}
# QDA for the model chosen by best subset selection/backward stepwise selection

set.seed(1)
qda_best <- qda(output~ sex+cp+exng+oldpeak+slp+caa+thall+trtbps_transformed+thalachh_transformed, 
                data = train2)
qda_best_pred <- predict(qda_best, test2)

table(qda_best_pred$class, test2$output)
```