---
title: "final project"
author: "Anna Barry"
date: "April 22, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
heart <- read.csv("heart.csv")
library(ggplot2)
library(tidyverse)
```


histogram of the age
```{r}
hist(heart$age, main = "Ages", xlab = "Age (in years)")
#lines(density(heart$age), col="red")
#lines(seq(25, 80, by=5), dnorm(seq(25, 80, by=5),
     # mean(heart$age), sd(heart$age)), col="blue")
curve(dnorm(x, mean=mean(heart$age), sd=sd(heart$age)), 
      col="darkblue", lwd=2, add=TRUE, yaxt="n")
```


cholestrol by age and sex, seems to be a lot less for sex = 0
```{r cars}
ggplot(data = heart, aes(x = age, y = chol)) + geom_point() + stat_smooth(method = lm) +
                    facet_wrap(.~sex) + labs(main = "chol v age for men and women") 
```

0 = female, 1 = male, more than twice as many men in this dataset
```{r}
heart %>% group_by(sex) %>% summarise( total = n())
```


output by age and sex, not what I would expect, 0 = female, 1 = male
```{r}
ggplot(data = heart, aes(x = age, y = output, color = as.factor(sex))) + geom_point() + 
   geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) + facet_wrap(.~sex)
```


type of angia by output
```{r}
heart %>% group_by(cp, output) %>% summarise(total = n())
```


covariance matrix with all predictors
```{r}
round(cov(heart), 2)

```

```{r}
round(var(heart), 2)

var(heart)
```


correlation amongst all the predictors, non seem to be super correlated
```{r}
round(cor(heart), 2)
```



angina type v age
```{r}
ggplot(data = heart, aes(x = age, y = cp)) + geom_point()
```



major number of vessels by age, there should not be 4. it says the range for caa is (0,3)
```{r}
ggplot(data = heart, aes(x = age, y = caa)) + geom_point()
```

fasting blood sugar by output
```{r}
heart %>% group_by(fbs, output) %>% summarise( total = n())
```


ecg by age, it looks like maybe as you age you may move up higher
```{r}
ggplot(data = heart, aes(x = age, y = restecg)) + geom_point()
```


resting ecg by output, maybe restecg = 1 is significant
```{r}
heart %>% group_by(output, restecg) %>% summarise(total = n())
```

seems like exng = 0 gives us more of a prob for having a heart attack. This makes sense, veins bring blood to heart, arteries take it away. The body will not let it enter the heart as fast.
```{r}
ggplot(data = heart, aes(x = age, y = output, color = as.factor(exng))) + 
      geom_point()   +  geom_smooth(method = "glm", 
                    method.args = list(family = "binomial"), se = FALSE)
```


maybe people just do not exercise
```{r}
heart %>% group_by(exng, restecg) %>% summarise(total = n())
```










